---
title: "introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(implant)

\usepackage{graphicx}
\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
\usepackage[authoryear, sort&compress]{natbib}
\usepackage{mathtools}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{verbatim}
\newcommand{\bdg}{{\bf g}}
\newcommand{\bdw}{{\bf w}}
\newcommand{\BB}{{\pmb B}}
\newcommand{\bdbeta}{{\pmb \beta}}
\newcommand{\bdmu}{{\pmb \mu}}
\newcommand{\T}{{\rm T}}
\newcommand{\bdOmega}{{\pmb \Omega}}
\def\wh{\widehat}

\author[1]{Ronghao Wang}
\author[2]{Yumou Qiu}
\author[1]{Yuzhen Zhou}
\author[3]{Yuhang Xu}
\author[4]{James C. Schnable}
\affil[1]{Department of Statistics, University of Nebraska-Lincoln, Lincoln, 68503,
USA}
\affil[2]{Department of Statistics, Iowa State University, Ames, 50011, USA}
\affil[3]{Department of Applied Statistics and Operations Research, Bowling Green State University, Bowling Green, 43403, USA}
\affil[4]{Center for Plant Science Innovation, Department of Agronomy and Horticulture, University of Nebraska-Lincoln, Lincoln, 68503, USA}

\begin{abstract}
    High-throughput phenotyping system has become more and more popular in plant science research. The data analysis for such a system typically involves two steps: plant feature extraction through image processing and statistical analysis for the extracted features. The current approach is to perform those two steps on different platforms.  We develop the package \textit{``implant''} in R for both robust feature extraction and functional data analysis, which is able to provide statistical inference for the plant traits directly from the input images. For image processing, the \textit{``implant''} package provides methods including thresholding, hidden Markov random field model, morphology operations and etc. For the growth curve analysis, this package can produce nonparametric curve fitting with its confidence region for plant growth. A functional ANOVA model to test for the treatment and genotype effects of the growth curve dynamics is also provided.
    %Functional ANOVA has many benefits in estimating the growth curve dynamics for plants, comparing to the point-wise ANOVA. This package provides functions to help users extract the phenotypical parameters from RGB images of plants and analyze them using Functional ANOVA model.
\end{abstract}



\begin{document}

\flushbottom
\maketitle

\thispagestyle{empty}
\newpage
\tableofcontents
\newpage
%\noindent Please note: Abbreviations should be introduced at the first mention in the main text – no abbreviations lists. Suggested structure of main text (not enforced) is provided below.
\section{Introduction}
This tutorial provides a guide of the $implant$ package for processing RGB images of  plants and using functional anova to analyze the growth dynamics for the plants. The whole procedure can be devided into two part, the data pre-processing part and functional analysis part. In the data pre-processing part, we first process the images of the plants of interest taken from the green house and using a series of steps to process these images to segment the bodies of these plants from their backgrounds. Then we can extract the phenotypical parameters such as size, height and width, etc., from the processed images. In the analysis part, we use the data extracted from the processed images in the first part, and use the functional anova approach to build models and have the corresponding analysis. 
\section{Image Processing}
In this part, we will use functions within this package to pre-process the RGB image data we have.
Our goal is to segment the object of interest (body of the plant) from its background, then we extract the phenotypical parameters from the segmented images.
%We will follow the following steps:\\
%1. Defining the Region of Interest;\\
%2. Segmenting the object of Interest;\\
%3. Morphorlogical Operations.\\\\
Figure1(a) from the Appendix is an image of a plant in a green house, and Figure1(b) is an image with exactly the same condition as Figure 1(a) except that it has no plant in the pot. Instead, a green strip was inserted into the pot.\\
\subsection{Defining the Region of Interest}
Using the package ``$png$", we can transform these two images into matrices of pixel intensities.
Let's name Figure 1(a) as $imageOriginal$ and Figure 1(b) as $image 2$. Since these two images all contain RGB channels, the figures should be read as three-dimensional arrays with the last dimensions is equal to 3 or 4. This depends on whether the images you use are RGB (3 channels) or RGBA (4 channels). In our case, the images we use has 4 channels, and we only want to consider the RGB channels. Therefore, we take the first three channels as our data matrices instead of 4. Then we can check the dimensions of these two matrices. For example, we'd like to find the dimensions for $imageOriginal$ for each channel:
\begin{verbatim}
    sizeOriginal = dim(imageOriginal)[c(1,2)]
    >2056 2454
\end{verbatim}
And we can see that in each channel, the matrix has 2056 rows and 2454 columns. Using these information, we can define our region of interest.\\\\
Using the function $ColorG$, we can find the green strip from $image2$ and change its color from green to the background color, white. Also we can find the vertical location of the top of the pot ( the location is the same as the location is the bottom part of the plant body in Figure 1(a).)  
\begin{verbatim}
    tempG = ColorG(image, rowThreshold = 0.002)
    lowerRowBound = tempG$lowb
\end{verbatim}
We may also reduce the sizes of these two images using function $sample$, and name the reduced original image and the reduced empty pot image as iamge1 and imageX1, respectively.
\begin{verbatim}
    image1 = sample(imageOriginal, RowSample = 2, ColSample = 2)
    imageX1 = sample(image2, RowSample = 2, ColSample = 2)
\end{verbatim}
And we can see that the sizes of these two images are both reduced from $2056 \times2454$ to $1028 \times1227$.\\\\
In order to obtain a well segmented result using thresholding method, we nned to identify the region of interest and change the color of the region that we are not interested in into black (Since we will segment the image in the next step and make the segmented image binary, which means coloring the background as black and the object of interest as white.). Using ``\textit{ColorB}" and ``$\textit{ColorG}$, we can identify the inner black bars and the border-top of the pot to obtain the region of interest for the plants. See the red rectangle in Figure \ref{fig:procedure} panel (b).
\begin{verbatim}
    #Identify the region of interest
    resultColor = ColorB(imageOriginal)
    ColBound = c(resultColor$lb, resultColor$rb)
    RI = matrix(1, sizeOrignal[1], sizeOrignal[2])
    RI[, c(1 : ColBound[1])] = 0
    RI[, c(ColBound[2] : sizeOrignal[2])] = 0
    RI[c(lowerRowBound : sizeOrignal[1]), ] = 0
    #Cut the image using the identified reigon of interest
    imageOriginalR = imageOriginal[, , 1]
    imageOriginalG = imageOriginal[, , 2]
    imageOriginalB = imageOriginal[, , 3]
    imageCutR = matrix(imageOriginalR[RI == 1],lowerRowBound - 1, 
                       ColBound[2] - ColBound[1] - 1)
    imageCutG = matrix(imageOriginalG[RI == 1],lowerRowBound - 1, 
                       ColBound[2] - ColBound[1] - 1)
    imageCutB = matrix(imageOriginalB[RI == 1],lowerRowBound - 1, 
                       ColBound[2] - ColBound[1] - 1)
    imageCut = array(c(imageCutR, imageCutG, imageCutB), dim = c(dim(imageCutR), 3))
\end{verbatim}
Notice that this identification strategy is for the dataset from the UNL greenhouse system only. Different systems need different strategies for locating the region of interest. 
\subsection{Image Segmentation using Thresholding Method}
We apply ``\textit{ColorG}" to imageX1 (Figure \ref{fig:procedure} (b)), we can change the color of the strip from green to white (the same color as the background). Then we can have a contrasted image (imageContrast) using the absolute value of the difference between image1 and imageX1G (see \ref{fig:procedure} (c) in the Appendix).
\begin{verbatim}
    tempG1 = ColorG(imageX1, rowThreshold = 0.002)
    imageX1G = tempG1$c
    imageContrast = abs(image1 - imageX1G)[, , 1 : 3]
\end{verbatim}
Function $imageBinary$ can help us segment the object of interest from the background and transform the segmented image into black and white, where the object of interest is colored as white and the background is colored as black. By defining the parameters of this function, we can have the segmented images for image1 and imageContrast, respectively. Then we can have a well segmented result by finding the intersection between the segmented image1 and the segmented imageContrast. See Figure \ref{fig:procedure} (d) - (f) in the Appendix.
\begin{verbatim}
    imageB1 = imageBinary(image1,  weight = c(-1, 2, -1), threshold1 = 30 / 255, 
    threshold2 = 0.02)
    imageB2 = imageBinary(imageContrast1,  weight = c(1, -2, 1), threshold1 = 0.7,
    threshold2 = 0)
    imageB = imageB1 * imageB2
\end{verbatim}
Note: An alternative method for segmentation we provide in this package is the Hidden Markov Random Field-EM Method. For details, please check the help documentation for the function \textit{``HMRF\_EM''}.
\subsection{Morphological Operations}
The next step is the morphological operations. Since there are still some unwanted noise, we would like to use the $erosion$ function to eliminate those noise. However, this action may cause the loose information of the segmented plant. Therefore, the function $dilation$ is needed as well. We follow the procedure: "Dilation-Erosion-Erosion-Dilation" and get the following result (see Figure \ref{fig:procedure} (g) in the Appendix).
\begin{verbatim}
    imageBD = dilation(imageB, mask = matrix(1, 5, 5))
    imageBDE = erosion( imageBD, mask = matrix(1, 5, 5) )
    imageBDEE = erosion( imageBDE, mask = matrix(1, 3, 3) )
    imageBDEED = dilation( imageBDEE, mask = matrix(1, 3, 3) )
\end{verbatim}
\subsection{Obtaining the Final Segmented Image}
From the "Defining the Region of Interest" step, we have a matrix $RI$ which has the same size as the original image, but the region that we are not interested in has been colored as black. Since our the size of the original image has been reduced, we need to reduce the size of $RI$ to the same size as $imageBDEED$ we got from the previous step. Then we cut the processed image (imageBDEED) and keep the region of interest(in order to remove the noise around the plant body).
\begin{verbatim}
    RI1 = sampleMatrix(RI, RowSample = 2, ColSample = 2)
    RI1row = sum(rowSums(RI1) > 0)
    RI1col = sum(colSums(RI1) > 0)
    imageBF = matrix(imageBDEED[RI1 == 1], RI1row, RI1col)
\end{verbatim}
$imageBF$ is the final segmented image we want. See Figure 1(i) in the Appendix.
\subsection{Extracting the Phenotypical Parameters}
Using the obtained segmented image, we can extract the information we want from its matrix. Firstly, we need to know how many millimeters that each pixel represents horizontally and vertically. The information we have is:
\begin{verbatim}
    Xsize (in mm): 1.54902601242065
    Ysize (in mm): 1.55327594280243
\end{verbatim}
which means each pixel from the image represents $1.549 mm$ in horizontal and $1.553 mm$ in vertical. Using this information, we can calcualte the size, height and weight from the pixel matrix of the segmented image.
\begin{verbatim}
    #a = 2, b = 2 since we have reduced the size by picking every two pixels 
    #in each row and each column
    extract(processed_image = imageBF, 
            Xsize = 1.54902601242065, Ysize = 1.55327594280243, a = 2, b = 2)$height
    extract(processed_image = imageBF, 
            Xsize = 1.54902601242065, Ysize = 1.55327594280243, a = 2, b = 2)$width
    extract(processed_image = imageBF, 
            Xsize = 1.54902601242065, Ysize = 1.55327594280243, a = 2, b = 2)$size
\end{verbatim}
Details can be found in the help documentation of this function.
\section{Functional ANOVA}
\subsection{Preparing for the Model Fitting}
In the previous part, we have extracted the phenotypical parameters we want from the processed images. In this part, we will use these parameters as our data to build Functional ANOVA models.\\\\
The example data we use in this tutorial is reduced from the data being described in the paper. The reduced data includes the size of $9$ plants with three genotypes ($78$, and $85$ and $129$) within $3$ different blocks. These plants were observed from $0^{th}$ day to the $44^{th}$ day. We have attached this example csv file in this package called \textit{data.csv}. We read the attached data and define the input terms as follows:
\begin{verbatim}
load_all()
data_new = read.csv(system.file("extdata", "data.csv",
                                 package = "implant", mustWork = TRUE)
Y.na.mat = data_new[,-c(1:3)]
time_interval = seq(from = 0, to = 44,by = 2)
X = data_new[,c(2:3)]
#This step is to factorize each factor
X1 = as.factor(unlist(X[,1]))
X2 = as.factor(unlist(X[,2]))
X = data.frame(X1,X2)
formula = toString("~ X[,1]+X[,2]")
\end{verbatim}
where $X$ is a matrix of explanatory variables, $Y.na.mat$ is the matrix of the extracted features, and $formula$ specifies the model we use, namely, Model (\ref{model}) in this example.
For both the covariate matrix $X$ and the response matrix $Y.na.mat$, each row gives the values for different plants. The columns of $X$ and $Y.na.mat$ correspond to explanatory variables and observation dates, respectively.
If a measurement of a plant is missing on a particular date, the value is filled by ``NA'' in the matrix.
More details can be found in the help documentation.
%By using the function $fanova$, we can estimate the model parameters. And we are able to have the estimated treatment effect and obtain the corresponding confidence intervals according to these estimated parameters got from $fanova$. Before using this function, there are a few steps we need to finish:\\
%1. Decide your model (including whether you want to have interactions or not.) This step is recommended using pointwise anova and check the significant factors before applying $fanova$.\\
%2. Create a Y matrix for the response variable. Each row should represent an observation and each column should represent an observing time point.\\
%3. Create an X matrix for the explanatory variables. Each column represents a factor you would like to involve in your model.\\
%4. Build a Time matrix for the observing time points. Each row of this matrix should be used for one observation and each column of this matrix should be used for the observed time from the original time point of this observation(i.e. the $ith$ row and the $jth$ column represents the $jth$ Day/Hour/Month/Year/Minute from the original observing time for the $ith$ observation. Each observation can have its own original observing time. The observing time can be regular or irregular.\\
%Note: if your observing time is regular, you can omit the Time matrix term.\\\\
%In this example, we would like to study how would the size of a plant be affected by its genotype effect and its block effect. In our experiment, we have 406 plants with 140 different genotypes in three blocks. Each genotype has 3 replicates. Within each block, a replicate of a genotype appears only once. Let's treat the block and genotype are two fixed effects in this example. Since the observing time for each plant is irregular, we have an irregular Time matrix T with 406 rows and 22 columns. Also, the dimension of Y matrix is $406 \times 22$ and the dimension of X matrix is $140*2$. For the model,we decide to use a model with no interactions between these two factors. Both factors are categorical variables.
\subsection{Model Fitting}
Based on the phenotypical parameters we obtained by processing images, we can study the relationship between the plant sizes and treatments, functional data models. Let $y_i(t)$ be the size of the $i^{th}$ plant measured at time $t$, where $i = 1,...,9$. We treat genotype and block as fixed effects. There are $3$ different genotypes and $3$ blocks in this study. We can use $\mathbf{G}_i= (G_{ik})_{k=2}^{3}$ as the categorical indicator of the $i^{th}$ plant genotype. Specifically, $G_{ik}$ is set to one if the plant has the $k^{th}$ genotype; otherwise, $G_{ik}=0$. For example, if the $i^{th}$ plant has the $2^{nd}$ genotype, then  $\mathbf{G}_i = (0, 1, 0, ..., 0)$. And $G_i$s being zeros represents the $1^{st}$ genotype, which is treated as the baseline. Similarly, $\mathbf{P_i} = (P_{ik})_{k=2}^3$ is defined to indicate the block that the $i^{th}$ plant belongs to, and the first block is set as the baseline. Our functional ANOVA model for analyzing the relationship between plant size and genotype, block is:
\begin{equation}
    y_i(t) = \mu(t)+ \sum_{k=2}^{3} G_{ik} g_k(t)+\sum_{k=2}^3 P_{ik}p_k(t) +\epsilon_i(t),
\end{equation}
where $\mu(t)$ is the growth function of the plant with the $1^{st}$ genotype (Genotype 78) from the $1^{st}$ block, $g_k(t)$s are the genotype effect functions, $p_k(t)$s are the fixed block effect functions, and the residuals $\epsilon_i(t)$s are modeled by independent random processes with mean zeros. The observation time points in this study are irregular.  Let $m_i$ be the number of  days that the $i^{th}$ plant was imaged, which could vary from plant to plant. Let $t_{ij}$ be the $j^{th}$ observation time of the $i^{th}$ plant, where $j = 1,...,m_i$. Then, our model can be described as:
\begin{equation}
    y_i(t_{ij}) = \mu(t_{ij})+ \sum_{k=2}^{3} G_{ik} g_k(t_{ij})+\sum_{k=2}^3 P_{ik}p_k(t_{ij}) +\epsilon_i(t_{ij}).
    \label{model}
\end{equation}
Since what we want is to study the relationship of the size and genotype effect, block effect, 
With all the input terms being prepared, we can fit the model using function \textit{fanova\_mean}:
\begin{verbatim}
  fit = fanova_mean(Y.na.mat, X = X, tt = time_interval, K.int = 6, order = 4,
                  interact = 0, p = 2, d0 = 0, d1 = 2, d2 = 2, 
                  lower = -10, upper = 15, formula = formula)
\end{verbatim}
%where $para$ is the a matrix of estimated parameters. Each of its columns represents the estimated parameter for a factor. In this example, its first column represents the estimated $\mu(t)$s. $\lambda$ is the estimated smooth parameter using GCV,  $bhat$ is a vector of all the estimated parameters, and $hatmatrix$ is the hat matrix of the model.
%\subsection{Treatment Effects}
%Function $est\_trt$ can be used for estimating the treatment effect of interest, using the estimated parameter matrix $para$ being obtained using $fanova$. For example, if we would like to know the estimated treatment effect of the $1st$ genotype and the $1st$ block, we can use:
%\begin{verbatim}
 % trt_g1b1 = est_trt(1,1,0, est_fun = para, X)  
%\end{verbatim}
%Similarly, the estimated treatment effect for the $2nd$ genotype in block $1$ is:
%\begin{verbatim}
%  trt_g2b1 = est_trt(2,1,0, est_fun = para, X)  
% plot(x = t.eval, y = trt_g1b1, type = "l", col = "1")
%lines(x = t.eval, y = trt_g2b1, type = "l", col = "2") 
%\end{verbatim}
\subsection{Statistical Inference}
By fitting the model, we can draw statistical inference,
such as parameter estimation, confidence bands, etc. For example, parameter estimation can be obtained by 
\begin{verbatim}
  para = fit$est_fun
\end{verbatim}
where $para$ is a $t$ by $q$ matrix. Here $t$ is the number of observation time points, and $q$ is the number of columns of the design matrix. In this case, $t = 23$, $q = 5$. The $j^{th}$ column represents the estimated parameters of the $j^{th}$ variable in the design matrix. For example, $\hat{\bf{\mu}}(t)$ can be obtained by:
\begin{verbatim}
  mu.t.hat = para[,1]
\end{verbatim}
Afterwards, we can test the significance of the treatment effects of interest by constructing the corresponding confidence regions with the function ``\textit{CI.trt.diff}" in the package. For example, we want to test the difference between block $2$ and block $1$: 
\begin{verbatim}
 blk2_1 = CI.trt.diff(fit = fit, j1 = 2, j2 = 1, alpha = 0.05)
\end{verbatim}
where $fit$ is the output from ``\textit{fanova\_mean}", and $j1$ and $j2$ specify the columns of the design matrix corresponds to the treatment of interest.
%\footnote{Add here.}
For example, the confidence region of the block effect function $p_3(\cdot)$ infers whether there is significant difference in plant size between block $1$ and block $3$, given the same plant genotype. Figure \ref{fig:ci} (a) shows the  $95\%$ confidence regions of $p_3$ from day $1$ to day $44$.
%\footnote{ronghao: add the maximum days}. 
We find that, compared to block $1$,  block $3$ does not have significant effect on the size of plants in the early stage of the growth until about the $15^{th}$ day.\\\\
In general, we can estimate a linear combination of treatment effects and its confidence regions by the following function:
\begin{verbatim}
 CI.trt(fit, L, alpha)
\end{verbatim}
where $L$ is a contrast vector under the model (\ref{model}) specifying the linear combination of the parameters of interest. 
This includes estimating the average growth curve of a particular genotype over all the blocks. For example,
\begin{verbatim}
 geno_78 = CI.trt(fit = fit, L = c(1,0,0,1/3,1/3), alpha = 0.05)
\end{verbatim}

%In order to calculate the point-wise confidence interval for $\hat{Y}(t)$, we need to know the variance of $\hat{Y}(t)$. 
%\begin{equation*}
%   Var(\hat{Y}(t)) = LS\Sigma_e S'L', 
%\end{equation*}
%where $L$ is the contrast vector, $S$ is the hatmatrix of the model, and $\Sigma_e$ is the covariance matrix of the residual term $\epsilon(t)$. Using the function $Sigma\_epsilon$, we can calculate the estimated sample covariance matrix of $\epsilon(t)$:
%\begin{verbatim}
 %  Cov = Sigma_epsilon(Y = Y, X = X, Time = T)
%\end{verbatim}
%Suppose we would like to know the confidence intervals for the $1st$ genotype in block $1$, and the confidence interval for the $1st$ genotype in block $2$. We construct our contrast vector $L$ and calculate the corresponding variance of $\hat{Y}(t)$ using function $Var\_yhat$:
%\begin{verbatim}
   % L = c(rep(1,K),rep(0,139*K),rep(0,2*K))
    %Var_yhat = Var_yhat(data_X = X,Time = T,plant_id = 1,total_time = 60,
     %                  K = 10, order = 4,d0 = 0,L = L , S = hatmatrix, Cov = Cov)
%\end{verbatim}
%Then we can calculate the confidence intervals using function $ci$:
\section{Appendix}
\begin{figure}[htp]
    \centering
        \begin{tabular}{cccc}
        \includegraphics[scale=0.08]{Main/Paper/image_plant.png}  &   			\includegraphics[scale=0.054]{Main/Paper/empty_pot.png} &
        \includegraphics[scale=0.08]{Main/Paper/image_contrast.png} &
        \includegraphics[scale=0.08]{Main/Paper/imageB_plant.png}\\
        (a) &   (b)  &      (c)  &      (d) \\
        \includegraphics[scale=0.081]{Main/Paper/imageBConstrast.png}   &   			\includegraphics[scale=0.081]{Main/Paper/imageBC_plant.png}  &
        \includegraphics[scale=0.081]{Main/Paper/imageBDEE_plant.png} &
        \includegraphics[scale=0.091]{Main/Paper/imageBFinalreduce111.png}\\
        (e) &   (f)  &     (g)       &  (h)
        \end{tabular}
    \caption{(a) Original plant image. (b) Original empty pot image; the red square is the identified region of interest by the functions \textit{``ColorB''} and \textit{``ColorG''}. (c) Contrast of (a) and (b). (d) Segmented image of (a) using DCT (e) Segmented image of (c) using DCT.  (f) Intersection of (d) and (e).  (g) Dilated-Eroded-Eroded-Dilated image of (f). (h) Final segmented image by identifying the region of interest.}
      \label{fig:procedure}
\end{figure}
%
\begin{figure}[htp]
    \centering
        \begin{tabular}{cccc}
        \includegraphics[height = 8cm, width = 8cm]{Main/Paper/blk2_ci.PNG}  &   			\includegraphics[height = 8cm, width = 8cm]{Main/Paper/gen_78_ci.PNG} \\
        (a) &   (b) 
        \end{tabular}
    \caption{(a) $95 \%$ confidence band of the coefficient of the difference between block 2 and block 1. (b) $95 \%$ confidence band of the estimated mean of genotype 78 averaging over three blocks.}
    \label{fig:ci}
\end{figure}
